\vfill World of Warcraft's New Way to Get You Hooked

### 2013-04-05 paper.tex
17. Proofreading. Missing words, transitions, better ways to say things, etc.

\subsubsection{Interests of the Software Engineer}
It is common sense that a software engineer strives to make working software. For World of Warcraft's software engineers, this success is measured by how many players enjoy new content and continue to play the game. Considerations are made for balance and gameplay as well as story. \cite{DevWatercooler} These engineers are also interested in keeping their jobs and continuing to make money. In order to keep their jobs, their interests must lie in getting players buy the software and then to come back to the game world often and play the game.

@@

First will be a purely numerical, Utilitarian approach to public good, `` ``Public Good'' By The Numbers''. The metric used will be the percentage of the population addicted to World of Warcraft. I find this argument to be a bit heartless and cold and so I will follow it with a section that shows that the public good is not only upheld, but improved, by World of Warcraft, `` ``Public Good'' Being Improved''. These improvements come from World of Warcraft helping some users overcome deficiencies in themselves, offering players powerful emotion experiences, and providing new or improved life skills.

@@

The most shocking statistic I found came from Dr. Maressa Orzack, founder and coordinator of Computer Addiction Services \cite{CompAddictionServices}, who cites that up to 40\% of World of Warcraft players could be addicted. \cite{FortyPercentAddicted} If everyone in the world played World of Warcraft, a total of over 7 billion people \cite{WorldPopulation}, and Dr. Ozark's prediction is correct, 2.8 billion people would be addicted. Based purely on the numbers that leaves 60\% of the world's population that does not have the mental disease. A utilitarian view in this case would be that World of Warcraft does not create mental disease in the majority of the world and therefore does not diminish mental or physical welfare of the public.

@@

\paragraph{Blizzard Gets Thank-You Letters\\}
During a panel at D.I.C.E. 2012 \cite{DICEInterview} that included Mike Epps, President of Epic Games, Ted Price, CEO and Founder of Insomniac Games, and Frank Pearce, current VP of product development and one of the three founders of Blizzard, Frank Pearce discussed a letter he received from a fan who had a disability and found comfort in World of Warcraft. He went on to say that Blizzard Entertainment is doing so many people a service that they actually receive letters thanking them for making such an immersive experience. \cite{DICEInterview}

\paragraph{World Of Warcraft Teaches Players Life Skills\\}
In a three year study of over 30,000 MMORPG users, over 50\% of users said that they learned valuable leadership skills that have helped them outside of the game. \cite{MotivationsAndDerviedExperiences}
Many features in World of Warcraft drive the players to socialize. \footnote{Having over two months of play time in-game with the last play session less than a week ago, I consider myself qualified to discuss the mechanics of World of Warcraft without citation.} The best gear and most prestigious titles are only available to users who group together to take on challenges that cannot be conquered by oneself. These challenges require intense coordination and communication as well as skilled players. Without grouping together, players have no chance of getting these items or attaining these statuses.
Guilds offer perks to their members through mechanics built into the game. As guilds grow in size and their members complete quests and dungeons, these guilds level up and offer a variety of rewards. Becoming part of a community of people is rewarded in-game.
\newline
\newline
Friendships and powerful experiences help us grow as individuals. World of Warcraft is improving the public good by improving the individuals that play it in ways that promote social skills and leadership skills. These skills benefit not only the player but also the people the player interacts with outside of the game. MMORPGs, like World of Warcraft, are providing their users with leadership skills, strong friendships, and powerful experiences that they take with them to interactions outside the game.

\subsubsection{Ethical Conclusion}
In order for Blizzard's actions to be ethical, its engineers must ``moderate the interests of themselves, Blizzard Entertainment, and the players of World of Warcraft with the public good.''

@@

Blizzard's software engineers are moderating all parties' interests. The public good is not only being upheld, but improved. Therefore, Blizzard Entertainment's actions are not only ethical in accordance with SE Code section 1.02, but can be seen as extraordinary, going beyond moderating their interests and improving the public good.

>>>>

\subsubsection{Interests of the Software Engineer}
It is common sense that a software engineer strives to make working software. For World of Warcraft's software engineers, this success is measured by how many players enjoy new content and continue to play the game. Considerations are made for balance and gameplay as well as story. \cite{DevWatercooler} These engineers are also interested in keeping their jobs and continuing to make money. In order to keep their jobs, their interests must lie in getting players to buy the software and then to come back to the game world often and play the game.

@@

First will be a purely numerical Utilitarian approach to public good, `` ``Public Good'' By The Numbers''. The metric used will be the percentage of the population addicted to World of Warcraft and thus the percentage of the ``public good'' that is being diminished. I find this argument to be a bit heartless and cold and so I will follow it with a section that shows that the public good is not only upheld, but improved, by World of Warcraft, `` ``Public Good'' Being Improved''. These improvements come from World of Warcraft helping some users overcome deficiencies in themselves, offering players powerful emotion experiences, and providing new or improved life skills.

@@

The most shocking statistic I found came from Dr. Maressa Orzack, founder and coordinator of Computer Addiction Services \cite{CompAddictionServices}, who cites that up to 40\% of World of Warcraft players could be addicted. \cite{FortyPercentAddicted} If everyone in the world played World of Warcraft, a total of over 7 billion people \cite{WorldPopulation}, and Dr. Ozark's prediction is correct, 2.8 billion people would be addicted. Based purely on the numbers that leaves 60\% of the world's population that does not have the mental disease. A utilitarian view in this case would be that World of Warcraft does not create mental disease in the majority of the world and therefore does not diminish the mental welfare of the public enough to be against the public good.

@@

\paragraph{Blizzard Gets Thank-You Letters\\}
During a panel at D.I.C.E. 2012 \cite{DICEInterview} that included Mike Epps, President of Epic Games, Ted Price, CEO and Founder of Insomniac Games, and Frank Pearce, current VP of product development and one of the three founders of Blizzard Entertainment, Frank Pearce discussed a letter he received from a fan who had a disability and found comfort in World of Warcraft. \cite{DICEInterview} He went on to say that Blizzard Entertainment is receives numerous letters thanking them for making such an immersive experience. \cite{DICEInterview}

\paragraph{World Of Warcraft Teaches Players Life Skills\\}
During this period of immersion, players are developing their personal skill sets. In a three year study of over 30,000 MMORPG users, over 50\% of users said that they learned valuable leadership skills that have helped them outside of the game. \cite{MotivationsAndDerviedExperiences}
Many features of World of Warcraft drive the players to socialize. \footnote{Having over two months of play time in-game with the last play session less than a week ago, I consider myself qualified to discuss the mechanics of World of Warcraft without citation.} The best gear and most prestigious titles are only available to users who group together to take on challenges that cannot be conquered by oneself. These challenges require intense coordination and communication as well as skilled players. Without grouping together, players have no chance of getting these items or attaining these statuses.
Guilds offer perks to their members through mechanics built into the game. As guilds grow in size and their members complete quests and dungeons, these guilds level up and offer a variety of rewards. Becoming part of a community of people is rewarded in-game.
\newline
\newline
Friendships and powerful experiences help us grow as individuals. World of Warcraft is improving the public good by improving the individuals that play it in ways that promote social skills and leadership skills. These skills benefit not only the player but also the people the player interacts with outside of the game. MMORPGs, like World of Warcraft, are providing their users with games that help the users improve themselves and their interactions with others.

\subsubsection{Ethical Conclusion}
In order for Blizzard's World of Warcraft to be ethical, its engineers must ``moderate the interests of themselves, Blizzard Entertainment, and the players of World of Warcraft with the public good.''

@@

Blizzard's software engineers are moderating all parties' interests. The public good is not only being upheld, but improved. Therefore, Blizzard Entertainment's World of Warcraft is not only ethical in accordance with SE Code section 1.02, but can be seen as extraordinary, going beyond just moderating their interests and improving the public good.
